ABS data cleaning 
=================
  
This script is used to load and clean raw demographic data from the 
Australian Bureau of Statistics (ABS). These data are primarily used for 
care cascade calculations but can be useful for Australian evaluation
models in general. The original raw data sets are stored 
in the ~/Evaluation\_Modelling/data/ABS/raw/ folder. The resulting cleaned
data sets are saved as data frames in csv files in the
~/Evaluation\_Modelling/data/ABS/ folder. A description of the ABS data sets 
and important data definitions are provided in each chunk. Each chunk can 
be run separately. 

Cleaned data is also stored/copied in specific project folder as required. 
In particular ~/Evaluation\_Modelling/project\_care\_cascades/data.

```{r Initialization}
# Restart R and set working directory to source file location

# Libraries for data manipulation
require(LeftysRpkg)
LoadLibrary(tidyverse)
LoadLibrary(lubridate)
LoadLibrary(readxl)

# Key folder paths
basePath <- file.path(dirname(getwd()))
absFolder <- file.path(basePath, "data", "ABS", "raw") # Folder with raw ABS data
cleanDataFolder <- file.path(basePath, "data", "ABS") # Where cleaned data goes

# Additional cleaned data folders for copies
cascadeDataFolder <- file.path("C:", "Users", "rgray", "UNSW",
  "Australian Care Cascades - Documents", "data")

# Source useful functions
source("TidyLongitudinal.R")

# Order of states used by ABS
states <- c("nsw", "vic", "qld", "sa", "wa", "tas", "nt", "act", "all")

# Primary script parameters
dataYear <- 2022 # Annual data from 2014

```

## Estimated resident population

```{r ERP}
# Process ERP data for each jurisdiction and overall

# Read in second sheet skipping most of the bumph
erpData <- read_excel(file.path(absFolder, 
  paste0("310104-ABS_ERP_State_Overall-", toString(dataYear), ".xlsx")),
  sheet = 2, skip = 9)

# Delete unnecessary columns (keeping the last nine) 
ncolumns <- ncol(erpData)
erpData <- erpData[, c(1, 20:ncolumns)]

# Convert first column to years 
erpData[, 1] <- year(erpData$`Series ID`)

# Rename columns
colnames(erpData) <- c("year", states)

# Remove unnecessary rows (keeping the final quarter data for each year)
# Starts in second quarter 1981 so selected ever third row. Then organize
# into long format with correct data types.
nrows <- nrow(erpData)
erpData <- erpData %>%
  slice(seq(3, nrows, by = 4)) %>%
  gather(state, erp, 2:10) %>%
  arrange(year) %>%
  mutate(year = as.numeric(year),
         state = as.character(state))

# Save the final cleaned data set 
write.csv(erpData, file = file.path(cleanDataFolder, 
  paste0("ABS_ERP_clean-", toString(dataYear) ,".csv")), 
  row.names = FALSE)

write.csv(erpData, file = file.path(cascadeDataFolder, 
  paste0("ABS_ERP_clean-", toString(dataYear) ,".csv")), 
  row.names = FALSE)
```

## Proportion of people who depart Australia - permantly or for more than
a year 

**Data:** ABS series 340102
**File:** 340102-ABS_Overseas_Departures-(dataYear).xls
**Definitions:**
- Permanent departures: Australian residents (including former settlers)
who on departure state on their passenger card that they are departing 
permanently.
- Long-term resident departures (LTRD): Australian residents who state
that they intend to stay abroad for 12 months or more (but not
permanently).
- Long-term visitor departures (LTVD): Overseas visitors departing after a
recorded stay of 12 months or more in Australia. 

Note: dataYear in the file refers to the last complete year in the data

```{r Departures}
# This script cleans the ABS data into the format we want. The raw data is
# always on the second sheet of the .xls data file.

# National ERP data ------------------------------------------------------

nationalERP <- filter(erpData, state == "all") %>%
  select(-state)

# Process overall departure data -----------------------------------------

# Read in second sheet skipping most of the bumph
if (dataYear < 2021) {
  departData <- read_excel(file.path(absFolder,
    paste0("340102-ABS_Overseas_Departures-", toString(dataYear), ".xls")),
                      sheet = 2, skip = 9)
} else {
  departData <- read_excel(file.path(absFolder,
    paste0("340102-ABS_Overseas_Departures-", toString(dataYear), ".xlsx")),
                      sheet = 2, skip = 9)
} 
# Delete unnecessary columns (keeping the first 4; quarter, permanent,
# long-term resident, long-term visitor) 
departData <- departData[, c(1:4)]

# Convert first column to years 
departData[, 1] <- year(departData$`Series ID`)

# Rename columns
colnames(departData) <- c("year", "permanent", "resident", "visitor")

# Group by year and sum to get total for each year
departData <- departData %>%
  group_by(year) %>%
  summarise(permanent = sum(permanent),
            resident = sum(resident),
            visitor = sum(visitor)) %>%
  mutate(year = as.numeric(year)) %>%  # Convert to correct data type
  ungroup() %>%
  inner_join(nationalERP, ., by = "year") %>% # Add overall ERP
  mutate(overall = permanent + resident + visitor,
         rate_permanent = permanent / erp, # Calculate departure rates
         rate_resident = resident / erp,
         rate_vistor = visitor / erp,
         overall_rate = (permanent + resident + visitor) / erp) 

# Save the final cleaned data set 
write.csv(departData, file = file.path(cleanDataFolder,
  paste0("ABS_departures_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

write.csv(departData, file = file.path(cascadeDataFolder,
  paste0("ABS_departures_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)
```

## Proportion of people who leave Australia by state and territory

The following chunk is used to estimate the number of people 
leaving Australia by state using the NOM departures estimate.

**Data:** ABS series 3412
**File:** 34120do007-ABS_Migration_NOM-(dataYear).xls
**ABS Explanatory notes:** 
http://www.abs.gov.au/ausstats/abs@.nsf/exnote/3412.0

NET OVERSEAS MIGRATION (NOM)

According to recommendations of the United Nations an international 
migrant is defined as "any person who changes his or her country of usual 
residence" (United Nations 1998). For the purposes of estimating NOM, and 
thereby Australia's official ERP counts, a person is regarded as a usual 
resident if they have been (or expect to be) residing in Australia for a 
period of 12 months or more over a 16 month period. As such, NOM and ERP 
estimates include all people, regardless of nationality, citizenship or 
legal status, who usually live in Australia, with the exception of foreign
diplomatic personnel and their families.

Conceptually, the term NOM is based on an international traveller's 
duration of stay being in or out of Australia for 12 months or more. It is
the difference between the number of incoming travellers who stay in 
Australia for 12 months or more and are added to the population (NOM 
arrivals) and the number of outgoing travellers who leave Australia for 12
months or more and are subtracted from the population (NOM departures). 
With the introduction of the improved methods for estimating NOM, this 12 
month period does not have to be continuous and is measured over a 
16-month reference period. Therefore, whether a traveller is in or out of 
the population is determined by their exact duration of stay in or away 
from Australia over the subsequent 16 months after arrival or departure.

Note: dataYear in the file refers to the ABS year end but the data is 
two years behind. First year is always 2004. 

```{r NOM}
# Process NOM departure data - Have to read in separate sheets

# Need to specify years first
if (dataYear %in% 2014:2016) {
  nomYear <- 2014
} else if (dataYear >= 2017) {
  nomYear <- 2017
}

years <- 2004:nomYear

nomDataYear <- min(dataYear, 2021)

# Initilize results data frame
nomData <- data.frame(year = as.numeric(),
                      state = as.numeric(),
                      arrivales = as.numeric(),
                      departures = as.numeric(),
                      nom = as.numeric())

# Loop through sheets extracting data we want
for (ii in seq(along = years) + 1) {
  
  # Read in sheet from file
  tempData <- read_excel(file.path(absFolder,                                  paste0("34120do007-ABS_Migration_NOM-", toString(nomDataYear) ,".xls")),
                         sheet = ii, skip = 5)
  
  # Replace annoying column names
  colnames(tempData) <- c("state", "grouping", "category", "arrivals",
                          "departures", "nom")
  
  # Select the data we want and do some general tidying up
  tempData <- tempData %>%
    select(-category) %>%
    filter(grouping == "Total") %>%
    mutate(state = states,
           year = rep(years[ii-1], length(states))) %>%
    select(-grouping, year, everything())
  
  # Append to final results
  nomData <- rbind(nomData, tempData)
}

# Order nomdata by state and add erpData and an overall departure rate
nomData <- arrange(nomData, year) %>%
  select(year, everything()) %>%
  inner_join(erpData, ., by = c("year", "state")) %>%
  mutate(depart_rate = departures / erp)

# Save the final cleaned data set 
write.csv(nomData, file = file.path(cleanDataFolder,
  paste0("ABS_migration_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

write.csv(nomData, file = file.path(cascadeDataFolder,
  paste0("ABS_migration_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```

## Proportion of people who move interstate

```{r interstate}

# Process arrivals read in second sheet ----------------------------------
if (dataYear < 2021) {
  arrData <- read_excel(file.path(absFolder,
    paste0("3101016a-ABS_Interstate_Arrivals-", 
      toString(dataYear), ".xls")), sheet = 2, skip = 9)
} else {
  arrData <- read_excel(file.path(absFolder,
    paste0("3101016a-ABS_Interstate_Arrivals-", 
      toString(dataYear), ".xlsx")), sheet = 2, skip = 9)
} 

# Convert first column to years 
arrData[, 1] <- year(arrData$`Series ID`)

# Convert colnames
colnames(arrData) <- c("year",states)

# Remove first three rows as they form an incomplete year
nrows <- nrow(arrData)
arrData <- arrData[4:nrows,]

numYears <- nrow(arrData)/4

# Sum the year data cumulative and then slice to select every fourth row 
# (corresponding to total for the year)

# First gather into long format
arrData <- arrData %>%
  gather(state, qtr_arrivals, 2:10) %>%
  group_by(year,state) %>%
  mutate(arrivals = cumsum(qtr_arrivals)) %>%
  ungroup() %>%
  slice(seq(4, numYears*8*4+4, by =4)) %>% # Exclude all; take every 4th row
  select(-qtr_arrivals)
 
# Process departures -----------------------------------------------------
if (dataYear < 2021) {
  depData <- read_excel(file.path(absFolder,
    paste0("3101016b-ABS_Interstate_Departures-", 
      toString(dataYear), ".xls")), sheet = 2, skip = 9)
} else {
  depData <- read_excel(file.path(absFolder,
    paste0("3101016b-ABS_Interstate_Departures-", 
      toString(dataYear), ".xlsx")), sheet = 2, skip = 9)
}

# Convert first column to years 
depData[, 1] <-  year(depData$`Series ID`)

# Convert colnames
colnames(depData) <- c("year",states)

# Remove first three rows as they form an incomplete year
nrows <- nrow(depData)
depData <- depData[4:nrows,]

# Sum the year data cumulative and then slice to select every fourth row 
# (corresponding to total for the year)

# First gather into long format
depData <- depData %>%
  gather(state, qtr_departures, 2:10) %>%
  group_by(year, state) %>%
  mutate(departures = cumsum(qtr_departures)) %>%
  ungroup() %>%
  slice(seq(4, numYears*8*4+4, by =4)) %>% # Exclude all; take every 4th row
  select(-qtr_departures)

# Join the arrivals and departures data frames
interstateData <- inner_join(arrData,depData, by=c("year", "state"))

# Save the final cleaned data set 
write.csv(interstateData, file = file.path(cleanDataFolder,
  paste0("ABS_interstate_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

write.csv(interstateData, file = file.path(cascadeDataFolder,
  paste0("ABS_interstate_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```

## Estimated resident population for each age and sex

This R chuck cleans the ERP data by sex and age provided by the ABS. It is
currently used in the Chlamydia incidence model. 

```{r group ERP}
# Process ERP for sex and age from second and third sheets

if (dataYear < 2021) {
  dataFile <- file.path(absFolder, paste0("3101059-ABS_ERP_Age_Sex-",
    toString(dataYear), ".xls"))
} else {
  dataFile <- file.path(absFolder, paste0("3101059-ABS_ERP_Age_Sex-",
    toString(dataYear), ".xlsx"))
}

erpData1 <- read_excel(dataFile, sheet = 2, skip = 9)

erpData2 <- read_excel(dataFile, sheet = 3, skip = 9)

# Combine data frames
erpasData <- cbind(erpData1, erpData2[, 2:ncol(erpData2)])

# Setup column names - 101 ages for males, females and all
ages <- as.character(0:100)
colnames(erpasData) <- c("year", paste("males-", ages, sep = ""),
                         paste("females-", ages, sep = ""),
                         paste("all-", ages, sep = ""))
 
# Convert first column to years 
erpasData[, 1] <- format(erpasData[, 1],"%Y")
erpasData$year <- as.numeric(erpasData$year)

# Save the final cleaned data set 
write.csv(erpasData, file = file.path(cleanDataFolder,
  paste0("ABS_ERP_sex_age_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

write.csv(erpasData, file = file.path(cascadeDataFolder,
  paste0("ABS_ERP_sex_age_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```

# Deprecated
#===========

The following chuck is no longer needed as we no longer use PharmDash data but 
it easily be run every year. 

```{r PharmDash ERP}
# Generate ERP for sub-populations in the ART data from the 10% 
# Prospection sample 
# Use variables: erpData and erpasData

prospectionYears <- 2013:dataYear

# Overall and state ERP
stateERP <- erpData %>% 
  filter(year %in% prospectionYears) %>%
  arrange(year, desc(erp)) %>%
  rename(population = state) 

# Sex and age ERP
asERP <- erpasData %>%
  filter(year %in% prospectionYears) %>%
  gather(2:ncol(.), key = "group", value = "erp") %>%
  arrange(year, group) %>%
  mutate(sex = str_split_fixed(.$group, "-", n = 2)[, 1], 
    age = str_split_fixed(.$group, "-", n = 2)[, 2]) %>%
  select(year, sex, age, erp) %>%
  mutate(age = as.integer(age)) %>%
  arrange(year, sex, age) %>%
  as_tibble()

sexERP <- asERP %>%
  group_by(year, sex) %>%
  summarise(erp = sum(erp)) %>% 
  rename(population = sex) %>%
  filter(population != "all") %>%
  mutate(population = str_replace(population, "s", "")) %>%
  ungroup()

ageERP20 <- asERP %>% 
  filter(sex == "all", age < 20) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "< 20") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP20_29 <- asERP %>% 
  filter(sex == "all", age %in% 20:29) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "20-29") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP30_39 <- asERP %>% 
  filter(sex == "all", age %in% 30:39) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "30-39") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP40_49 <- asERP %>% 
  filter(sex == "all", age %in% 40:49) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "40-49") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP50_59 <- asERP %>% 
  filter(sex == "all", age %in% 50:59) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "50-59") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP60_69 <- asERP %>% 
  filter(sex == "all", age %in% 60:69) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "60-69") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP70 <- asERP %>% 
  filter(sex == "all", age > 70) %>%
  group_by(year) %>%
  summarise(erp = sum(erp)) %>%
  mutate(age = "70+") %>%
  select(year, age, erp) %>%
  ungroup()

ageERP <- bind_rows(ageERP20, ageERP20_29, ageERP30_39, ageERP40_49,
  ageERP50_59, ageERP60_69, ageERP70) %>%
  arrange(year, age) %>% 
  rename(population = age)

# Combine all into file data frame
prospectionERP <- bind_rows(stateERP, sexERP, ageERP) %>%
  arrange(year) %>%
  select(population, year, erp)

# Save the final cleaned data set 
write.csv(prospectionERP, file = file.path(cleanDataFolder,
  paste0("ABS_population_sizes_PharmDash-", dataYear, ".csv")), 
  row.names = FALSE)

write.csv(prospectionERP, file = file.path(cascadeDataFolder,
  paste0("ABS_population_sizes_PharmDash-", dataYear, ".csv")), 
  row.names = FALSE)

```

# Once off data processing
# ========================

The following chunks process once off data files from the ABS. 

```{r Interstate migration by age and sex}
# This chuck reads in the saved csv file downloaded from ABS.Stat with the 
# data for Interstate migration: Arrivals, departures and net, 
# State/territory, Age and sex - Calendar years, 1997 onwards. 

# Read in and do basic tidying of downloaded file
interMigrationData <- read_csv(file.path(absFolder,
  paste0("3101-ABSSTAT_Interstate_movement_age_sex-1997-2016.csv"))) %>%
  select(-MEASURE, -AGE, -SEX_ABS, -ASGS_2011, -FREQUENCY, -Frequency, 
    -TIME, -Flags, -`Flag Codes`) %>%
  # Rename columns
  rename(type = `Interstate Migration`, sex = Sex, age = Age, 
    region = Region, year = Time, value = Value) %>%
  select(year, region, sex, age, type, value) %>%
  mutate(region = tolower(region), sex = tolower(sex), 
    type = tolower(type)) %>%
  # Tidy up strings
  mutate(age = str_replace_all(age, " ", ""),
    type = str_sub(type, 5, -1),
    sex = ifelse(sex == "persons", "all", sex)) %>%
  mutate(sex = ifelse(sex == "males", "male", sex),
    sex = ifelse(sex == "females", "female", sex)) %>%
  mutate(age = str_c("a", str_replace_all(age, "-", "_"))) %>%
  mutate(age = ifelse(age == "a75andover", "a75+", age)) %>%
  mutate(age = ifelse(age == "aAllages", "all", age)) %>%
  # Rename all the regions
  mutate(region = ifelse(region == "australia", "all", region)) %>%
  mutate(region = ifelse(region == "australian capital territory", "act",
    region)) %>%
  mutate(region = ifelse(region == "new south wales", "nsw", region)) %>%
  mutate(region = ifelse(region == "northern territory", "nt", region)) %>%
  mutate(region = ifelse(region == "other territories", "other", region)) %>%
  mutate(region = ifelse(region == "queensland", "qld", region)) %>%
  mutate(region = ifelse(region == "south australia", "sa", region)) %>%
  mutate(region = ifelse(region == "tasmania", "tas", region)) %>%
  mutate(region = ifelse(region == "victoria", "vic", region)) %>%
  mutate(region = ifelse(region == "western australia", "wa", region)) %>%
  # Only keep departures and arrivals and remove other
  filter(type != "interstate", region != "other")

# Save the final cleaned data set 
write.csv(interMigrationData, file = file.path(cleanDataFolder,
  paste0("ABS_interstate_age_sex_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

write.csv(interMigrationData, file = file.path(cascadeDataFolder,
  paste0("ABS_interstate_age_sex_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```

## Proportion of people who move between SA3 regions

```{r Interegion movement}
# This chuck reads in the saved csv files downloaded from ABS.Stat with the 
# data for migration bewteen SA3 level geographic areas (Arrivals, 
# departures, region, age group and sex - financial years 2006-07 to 
# 2014-15) and ERP data for each SA3 region. 

# To obtain data movement data go to series 3218.0 Regional Population
# Growth, Australia 2017-18
# (https://www.abs.gov.au/AUSSTATS/abs@.nsf/mf/3218.0) and enter
# ABS.Stat. Select Regional Internal Migration (Regional Internal Migration 
# Estimates by Region (SA3 and above), Age and Sex, 2006-07 to 2015-16   : 
# RIME by Region). To download the data requires selecting the SA3 regional 
# level from Customize and downloading the persons, males, and females data 
# separately to three CSV files: 
# 3218-ABSSTAT_regional_movement_age_sex-2007-2015-persons.csv, 
# 3218-ABSSTAT_regional_movement_age_sex-2007-2015-males.csv, and
# 3218-ABSSTAT_regional_movement_age_sex-2007-2015-females.csv.
# This is because the data is too big to download as a single file. This data # is merged and cleaned below. 
# 
# To obtain the ERP data go to series 3218.0 Regional Population
# Growth, Australia 2017-18
# (https://www.abs.gov.au/AUSSTATS/abs@.nsf/mf/3218.0) and enter
# ABS.Stat. Select Estimated Resident Population (ERP by SA2 (ASGS 2011), 
# Age and Sex, # 2001 to 2016). To download the data requires selecting the 
# SA3 regional level from Customize and downloading the persons, males, and 
# females data separately to three CSV files: 
# 3218-ABSSTAT_ERP_Regional_Age_Sex-2001-2016-persons.csv, 
# 3218-ABSSTAT_ERP_Regional_Age_Sex-2001-2016-males.csv, and
# 3218-ABSSTAT_ERP_Regional_Age_Sex-2001-2016-females.csv.
# This is because the data is too big to download as a single file. This 
# data is merged and cleaned below. 

## First sort out population movement -------------------------------------

# Read in and do basic tidying of downloaded movement files
regionMigrationDataPersons <- read_csv(file.path(absFolder,
  paste0("3218-ABSSTAT_regional_movement_age_sex-2007-2015-persons.csv")),
  col_types = cols(
  RIM = col_integer(),
  `Migration Type` = col_character(),
  AGE = col_character(),
  `Age Group` = col_character(),
  SEX = col_integer(),
  Sex = col_character(),
  REGIONTYPE = col_character(),
  `Region Type` = col_character(),
  ASGS_2011_SA34_GCCSA_STE = col_character(),
  Region = col_character(),
  FREQUENCY = col_character(),
  Frequency = col_character(),
  TIME_FY = col_integer(),
  Time = col_character(),
  Value = col_integer(),
  `Flag Codes` = col_character(),
  Flags = col_character())) %>%
  select(`Migration Type`, `Age Group`, Sex, REGIONTYPE, Region, TIME_FY,
    Value) 

regionMigrationDataMales <- read_csv(file.path(absFolder,
  paste0("3218-ABSSTAT_regional_movement_age_sex-2007-2015-males.csv")),
  col_types = cols(
  RIM = col_integer(),
  `Migration Type` = col_character(),
  AGE = col_character(),
  `Age Group` = col_character(),
  SEX = col_integer(),
  Sex = col_character(),
  REGIONTYPE = col_character(),
  `Region Type` = col_character(),
  ASGS_2011_SA34_GCCSA_STE = col_character(),
  Region = col_character(),
  FREQUENCY = col_character(),
  Frequency = col_character(),
  TIME_FY = col_integer(),
  Time = col_character(),
  Value = col_integer(),
  `Flag Codes` = col_character(),
  Flags = col_character())) %>%
  select(`Migration Type`, `Age Group`, Sex, REGIONTYPE, Region, TIME_FY,
    Value) 

regionMigrationDataFemales <- read_csv(file.path(absFolder,
  paste0("3218-ABSSTAT_regional_movement_age_sex-2007-2015-females.csv")),
  col_types = cols(
  RIM = col_integer(),
  `Migration Type` = col_character(),
  AGE = col_character(),
  `Age Group` = col_character(),
  SEX = col_integer(),
  Sex = col_character(),
  REGIONTYPE = col_character(),
  `Region Type` = col_character(),
  ASGS_2011_SA34_GCCSA_STE = col_character(),
  Region = col_character(),
  FREQUENCY = col_character(),
  Frequency = col_character(),
  TIME_FY = col_integer(),
  Time = col_character(),
  Value = col_integer(),
  `Flag Codes` = col_character(),
  Flags = col_character())) %>%
  select(`Migration Type`, `Age Group`, Sex, REGIONTYPE, Region, TIME_FY,
    Value) 

# Bind the data and clean
regionMigrationData <- bind_rows(regionMigrationDataPersons, 
  regionMigrationDataMales, regionMigrationDataFemales) %>%
  # Only keep for SA3 level
  filter(REGIONTYPE == "SA3") %>%
  # Rename columns
  rename(type = `Migration Type`, age = `Age Group`, sex = Sex, 
    region = Region, year = TIME_FY, value = Value) %>%
  select(year, region, sex, age, type, value) %>%
  mutate(sex = tolower(sex), 
    type = tolower(type), age = tolower(age)) %>%
 # Tidy up strings
  mutate(age = str_replace_all(age, " ", ""),
    type = str_sub(type, 22, -1),
    sex = ifelse(sex == "persons", "all", sex)) %>%
  mutate(sex = ifelse(sex == "males", "male", sex),
    sex = ifelse(sex == "females", "female", sex)) %>%
  mutate(age = str_c("a", str_replace_all(age, "-", "_"))) %>%
  mutate(age = ifelse(age == "a65andover", "a65+", age)) %>%
  mutate(age = ifelse(age == "aallages", "all", age)) %>%
  # Spread type and drop net
  spread(type, value) %>%
  select(-net)

## Second sort out ERP ----------------------------------------------------

# Read in and do basic tidying of downloaded ERP files
regionErpDataPersons <- read_csv(file.path(absFolder,
  paste0("3218-ABSSTAT_ERP_Regional_Age_Sex-2001-2016-persons.csv")),
  col_types = cols(
    MEASURE = col_character(),
    Measure = col_character(),
    SEX_ABS = col_integer(),
    Sex = col_character(),
    AGE = col_character(),
    Age = col_character(),
    REGIONTYPE = col_character(),
    `Region Type` = col_character(),
    ASGS_2011 = col_integer(),
    Region = col_character(),
    FREQUENCY = col_character(),
    Frequency = col_character(),
    TIME = col_integer(),
    Time = col_integer(),
    Value = col_integer(),
    `Flag Codes` = col_character(),
    Flags = col_character()
  )) 

regionErpDataMales <- read_csv(file.path(absFolder,
  paste0("3218-ABSSTAT_ERP_Regional_Age_Sex-2001-2016-males.csv")),
  col_types = cols(
    MEASURE = col_character(),
    Measure = col_character(),
    SEX_ABS = col_integer(),
    Sex = col_character(),
    AGE = col_character(),
    Age = col_character(),
    REGIONTYPE = col_character(),
    `Region Type` = col_character(),
    ASGS_2011 = col_integer(),
    Region = col_character(),
    FREQUENCY = col_character(),
    Frequency = col_character(),
    TIME = col_integer(),
    Time = col_integer(),
    Value = col_integer(),
    `Flag Codes` = col_character(),
    Flags = col_character()
  ))

regionErpDataFemales <- read_csv(file.path(absFolder,
  paste0("3218-ABSSTAT_ERP_Regional_Age_Sex-2001-2016-females.csv")),
  col_types = cols(
    MEASURE = col_character(),
    Measure = col_character(),
    SEX_ABS = col_integer(),
    Sex = col_character(),
    AGE = col_character(),
    Age = col_character(),
    REGIONTYPE = col_character(),
    `Region Type` = col_character(),
    ASGS_2011 = col_integer(),
    Region = col_character(),
    FREQUENCY = col_character(),
    Frequency = col_character(),
    TIME = col_integer(),
    Time = col_integer(),
    Value = col_integer(),
    `Flag Codes` = col_character(),
    Flags = col_character()
  ))

# Bind the data and clean
regionErpData <- bind_rows(regionErpDataPersons, regionErpDataMales,
  regionErpDataFemales) %>%
  select(Sex, Age, REGIONTYPE, Region, Time, Value) %>%
  # Only keep for SA3 level
  filter(REGIONTYPE == "SA3") %>%
  # Rename columns
  rename(age = Age, sex = Sex, 
    region = Region, year = Time, erp = Value) %>%
  select(year, region, sex, age, erp) %>%
  mutate(sex = tolower(sex), 
    age = tolower(age)) %>%
  # Tidy up strings
  mutate(age = str_replace_all(age, " ", ""),
    sex = ifelse(sex == "persons", "all", sex)) %>%
  mutate(sex = ifelse(sex == "males", "male", sex),
    sex = ifelse(sex == "females", "female", sex)) %>%
  mutate(age = str_c("a", str_replace_all(age, "-", "_"))) %>%
  mutate(age = ifelse(age == "a85andover", "a85+", age)) %>%
  mutate(age = ifelse(age == "aallages", "all", age))

# Group ERP into movement age bins: "a0_14", "a15_24", "a25_44", "a45_64"
# "a65+", "all" 

regionErpData <- regionErpData %>%
  mutate(agebin = ifelse(age %in% c("a0_4", "a5_9", "a10_14"), "a0_14", 
    age)) %>%
  mutate(agebin = ifelse(age %in% c("a15_19", "a20_24"), "a15_24", 
    agebin)) %>%
  mutate(agebin = ifelse(age %in% c("a25_29", "a30_34", "a35_39", 
    "a40_44"), "a25_44", agebin)) %>%
  mutate(agebin = ifelse(age %in% c("a45_49", "a50_54", "a55_59", 
    "a60_64"), "a45_64", agebin)) %>%
  mutate(agebin = ifelse(age %in% c("a65_69", "a70_74", "a75_79", "a80_84", 
    "a85+"), "a65+", agebin))

regionErpData <- regionErpData %>%
  select(-age) %>%
  group_by(year, region, sex, agebin) %>%
  summarise(erp = sum(erp)) %>%
  ungroup() %>%
  rename(age = agebin)

# Add erp data to movement data -------------------------------------------
regionMigrationData <- regionMigrationData %>% 
  left_join(regionErpData, by = c("year", "region", "sex", "age"))

## Save the final cleaned data set ---------------------------------------- 
write.csv(regionMigrationData, file = file.path(cleanDataFolder,
  paste0("ABS_interregion_age_sex_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

write.csv(regionMigrationData, file = file.path(cascadeDataFolder,
  paste0("ABS_interregion_age_sex_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```

## Estimated death rates

```{r Overall deathrate}
# Overall death rate data is available from 1971 to 2014 by single year
# age. Only extract overall here.
deathDataAll <- read_excel(file.path(absFolder,
  paste0("33020-ABS_Overall_Death_Rates-1971_2014.xls")), skip = 6) %>%
  slice(2) %>%
  select(-Time) %>%
  gather("year", "deathrate")

# Save the final cleaned data set 
write.csv(deathDataAll, file = file.path(cleanDataFolder,
  "ABS_deathrate_all_clean-2014.csv"), 
  row.names = FALSE)

```

This R chuck cleans the ABS age-specific death rate data for males and 
females. This is used for estimating the proportion of PLDHIV in each five
year age bin. 

```{r deathrates}
# Currently date is only available up to 2014. Need to loop through each 
# sheet to get data for each state. Crude death rate per 1000 is converted
# to a proportion. 

# Need to specify years first
years <- 2004:2014

# Initilize results data frame
deathData <- data.frame(year = as.numeric(),
                      state = as.character(),
                      age = as.character(),
                      gender = as.character(),
                      deathrate = as.numeric())

# Loop through sheets extracting data we want
for (ii in 2:(length(states) + 1)) {
  
  # Read in sheet from file
  tempData <- read_excel(file.path(absFolder,           
      "33020do002_ABS_Deathrate_2014.xls"), sheet = ii, skip = 4)
  
  # Extract the overall deathrates ---------------------------------------
  tempDataAll <- tempData %>%
    slice(8:10) %>%
    mutate(state = states[ii - 1]) %>%
    select(state, everything())
  
  # Rename columns
  colnames(tempDataAll) <- c("state", "population", as.character(years))
  
  # Reorganize
  tempDataAll <- tempDataAll %>%
    gather("year", "deathrate", 3:ncol(tempDataAll)) %>%
    mutate(year = as.numeric(year)) %>%
    mutate(age = "all",
           gender = tolower(population),
           deathrate = as.numeric(deathrate)) %>%
    select(year, state, age, gender, deathrate)
  
  # Replae persons with all
  tempDataAll[tempDataAll$gender == "persons", ]$gender <- "all"
  
  # Now sort out male and female age groups ------------------------------        
  tempDataMales <- tempData %>%
    slice(17:35) %>%
    mutate(state = states[ii - 1]) %>%
    select(state, everything())
  
  # Rename columns
  colnames(tempDataMales) <- c("state", "age", as.character(years))
  
  # Reorganize
  tempDataMales <- tempDataMales %>%
    gather("year", "deathrate", 3:ncol(tempDataMales)) %>%
    mutate(year = as.numeric(year)) %>%
    mutate(gender = "males",
           deathrate = as.numeric(deathrate)) %>%
    select(year, state, age, gender, deathrate)

  # Females
  tempDataFemales <- tempData %>%
    slice(17:35) %>%
    mutate(state = states[ii - 1]) %>%
    select(state, everything())
  
  # Rename columns
  colnames(tempDataFemales) <- c("state", "age", as.character(years))
  
  # Reorganize
  tempDataFemales <- tempDataFemales %>%
    gather("year", "deathrate", 3:ncol(tempDataFemales)) %>%
    mutate(year = as.numeric(year)) %>%
    mutate(gender = "females",
           deathrate = as.numeric(deathrate)) %>%
    select(year, state, age, gender, deathrate)
  
  # Merge all the data into one data frame
  deathData <- bind_rows(deathData, tempDataAll, tempDataMales,
                         tempDataFemales) %>% tbl_df
  
}

# Tidy up the final data frame converting "85 and over" to 85+
deathData[deathData$age == "85 and over", ]$age <- "85+"

# Save the final cleaned data set 
write.csv(deathData, file = file.path(cleanDataFolder,
  "ABS_deathrate_age_state_clean-2014.csv"), 
  row.names = FALSE)

```

