ABS data cleaning 
=================
  
This script is used to load and clean raw demographic data from the 
Australian Burea of Statistica (ABS). These data are primarily used for 
care cascade calculations but can be useful for Australian evaluation
models in general. The original raw data sets are stored 
in the ~/Evaluation\_Modelling/data/ABS/ folder. The resulting cleaned
data sets are saved as dataframes in csv files in the
~/Evaluation\_Modelling/data/ folder. A description of the ABS data sets 
and important data definitions are provided in each section. 

```{r initialization}
# Clear workspace
rm(list=ls()) 

# Libraries for data manipulation
require(readxl)
require(dplyr)
require(tidyr)
require(utils)

# Key folder paths
Rpath <- file.path(path.expand("~"), "R") 
basePath <- file.path(path.expand("~"), "Research",
                      "!Evaluation_Modelling")
absFolder <- file.path(basePath,"data","ABS") # Folder with raw ABS data
cleanDataFolder <- file.path(basePath,"data")

# Source useful functions
source(file.path(Rpath, "TidyLongitudinal.R"), echo=TRUE)

# Set working directory
setwd(file.path(basePath,"code"))

# Order of states used by ABS
states <- c("nsw","vic","qld","sa","wa","tas","nt","act","all")

# Primary script parameters
dataYear <- 2015

```

## Estimated resident population

```{r ERP}
# Process ERP data for each jurisdiction and overall

# Read in second sheet skipping most of the bumph
erpData <- read_excel(file.path(absFolder, 
  paste0("310104-ABS_ERP_State_Overall-", toString(dataYear), ".xlsx")),
  sheet = 2, skip = 9)

# Delete unnecessary columns (keeping the last nine) 
ncolumns <- ncol(erpData)
erpData <- erpData[, c(1, 20:ncolumns)]

# Convert first column to years 
erpData[, 1] <- format(erpData[, 1], "%Y")

# Rename columns
colnames(erpData) <- c("year", states)

# Remove unnecessary rows (keeping the final quarter data for each year)
# Starts in second quarter 1981 so selected ever third row. Then organize
# into long format with correct data types.
nrows <- nrow(erpData)
erpData <- erpData %>%
  slice(seq(3, nrows, by = 4)) %>%
  gather(state, erp, 2:10) %>%
  arrange(year) %>%
  mutate(year = as.numeric(year),
         state = as.character(state))

# Save the final cleaned data set 
write.csv(erpData, file = file.path(cleanDataFolder, 
  paste0("ABS_ERP_clean-", toString(dataYear) ,".csv")), 
  row.names = FALSE)
```

## Proportion of people who depart Australia - permantly or for more than
a year 

**Data:** ABS series 340102
**File:** 340102-ABS_Overseas_Departures-(dataYear).xls
**Definitions:**
- Permanent departures: Australian residents (including former settlers)
who on departure state on their passenger card that they are departing 
permanently.
- Long-term resident departures (LTRD): Australian residents who state
that they intend to stay abroad for 12 months or more (but not
permanently).
- Long-term visitor departures (LTVD): Overseas visitors departing after a
recorded stay of 12 months or more in Australia. 

Note: dataYear in the file refers to the last complete year in the data

```{r Departures}
# This script cleans the ABS data into the format we want. The raw data is
# always on the second sheet of the .xls data file.

# National ERP data ------------------------------------------------------

nationalERP <- filter(erpData, state == "all") %>%
  select(-state)

# Process overall departure data -----------------------------------------

# Read in second sheet skipping most of the bumph
departData <- read_excel(file.path(absFolder,
  paste0("340102-ABS_Overseas_Departures-", toString(dataYear), ".xls")),
                      sheet = 2, skip = 9)

# Delete unnecessary columns (keeping the first 4; quarter, permanent,
# long-term resident, long-term visitor) 
departData <- departData[, c(1:4)]

# Convert first column to years 
departData[,1] <- format(departData[, 1], "%Y")

# Rename columns
colnames(departData) <- c("year", "permanent", "resident", "visitor")

# Group by year and sum to get total for each year
departData <- departData %>%
  group_by(year) %>%
  summarise(permanent = sum(permanent),
            resident = sum(resident),
            visitor = sum(visitor)) %>%
  mutate(year = as.numeric(year)) %>%  # Convert to correct data type
  ungroup() %>%
  inner_join(nationalERP, ., by = "year") %>% # Add overall ERP
  mutate(overall = permanent + resident + visitor,
         rate_permanent = permanent / erp, # Calculate departure rates
         rate_resident = resident / erp,
         rate_vistor = visitor / erp,
         overall_rate = (permanent + resident + visitor) / erp) 

# Save the final cleaned data set 
write.csv(departData, file = file.path(cleanDataFolder,
  paste0("ABS_departures_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```

## Proportion of people who leave Australia by state and territory

The following chunk is used to estimate the number of people 
leaving Australia by state using the NOM departures estimate.

**Data:** ABS series 3412
**File:** 34120do007-ABS_Migration_NOM-(dataYear).xls
**ABS Explanatory notes:** 
http://www.abs.gov.au/ausstats/abs@.nsf/exnote/3412.0

NET OVERSEAS MIGRATION (NOM)

According to recommendations of the United Nations an international 
migrant is defined as "any person who changes his or her country of usual 
residence" (United Nations 1998). For the purposes of estimating NOM, and 
thereby Australia's official ERP counts, a person is regarded as a usual 
resident if they have been (or expect to be) residing in Australia for a 
period of 12 months or more over a 16 month period. As such, NOM and ERP 
estimates include all people, regardless of nationality, citizenship or 
legal status, who usually live in Australia, with the exception of foreign
diplomatic personnel and their families.

Conceptually, the term NOM is based on an international traveller's 
duration of stay being in or out of Australia for 12 months or more. It is
the difference between the number of incoming travellers who stay in 
Australia for 12 months or more and are added to the population (NOM 
arrivals) and the number of outgoing travellers who leave Australia for 12
months or more and are subtracted from the population (NOM departures). 
With the introduction of the improved methods for estimating NOM, this 12 
month period does not have to be continuous and is measured over a 
16-month reference period. Therefore, whether a traveller is in or out of 
the population is determined by their exact duration of stay in or away 
from Australia over the subsequent 16 months after arrival or departure.

Note: dataYear in the file refers to the ABS year end but the data is 
two years behind. First year is always 2004. 

```{r NOM}
# Process NOM departure data - Have to read in separate sheets

# Need to specify years first
years <- 2004:(dataYear-2)

# Initilize results data frame
nomData <- data.frame(year = as.numeric(),
                      state = as.numeric(),
                      arrivales = as.numeric(),
                      departures = as.numeric(),
                      nom = as.numeric())

# Loop through sheets extracting data we want
for (ii in seq(along = years) + 1) {
  
  # Read in sheet from file
  tempData <- read_excel(file.path(absFolder,                                  paste0("34120do007-ABS_Migration_NOM-", toString(dataYear) ,".xls")),
                         sheet = ii, skip = 5)
  
  # Replace annoying column names
  colnames(tempData) <- c("state", "grouping", "category", "arrivals",
                          "departures", "nom")
  
  # Select the data we want and do some general tidying up
  tempData <- tempData %>%
    select(-category) %>%
    filter(grouping == "Total") %>%
    mutate(state = states,
           year = rep(years[ii-1], length(states))) %>%
    select(-grouping, year, everything())
  
  # Append to final results
  nomData <- rbind(nomData, tempData)
}

# Order nomdata by state and add erpData and an overall departure rate
nomData <- arrange(nomData, year) %>%
  select(year, everything()) %>%
  inner_join(erpData, ., by = c("year", "state")) %>%
  mutate(depart_rate = departures / erp)

# Save the final cleaned data set 
write.csv(nomData, file = file.path(cleanDataFolder,
  paste0("ABS_migration_clean-", toString(dataYear), ".csv")), 
  row.names = FALSE)

```



## Proportion of people who move interstate

```{r interstate}

### Process arrivals read in second sheet
arrdata <- read_excel(file.path(absFolder,"3101016a-ABS_Interstate_Arrivals.xls"),
                      sheet = 2, skip = 9)

# Convert first column to years 
arrdata[,1] <- format(arrdata[,1],"%Y")

# Convert colnames
colnames(arrdata) <- c("year",states)

# Remove first three rows as they form an incomplete year
nrows <- nrow(arrdata)
arrdata <- arrdata[4:nrows,]

# Sum the year data cumulative and then slice to select every fourth row 
# (corresponding to total for the year)

# First gather into long format
arrdata <- gather(arrdata, state,qtr_arrivals,2:10) 

# Now do the processing
arrdata <- arrdata %>%
  group_by(year,state) %>%
  mutate(arrivals = cumsum(qtr_arrivals)) %>%
  ungroup() %>%
  slice(seq(4,1188, by =4)) %>%
  select(-qtr_arrivals)
 

### Process departures
depdata <- read_excel(file.path(absFolder,"3101016b-ABS_Interstate_Departures.xls"),
                      sheet = 2, skip = 9)

# Convert first column to years 
depdata[,1] <- format(depdata[,1],"%Y")

# Convert colnames
colnames(depdata) <- c("year",states)

# Remove first three rows as they form an incomplete year
nrows <- nrow(depdata)
depdata <- depdata[4:nrows,]

# Sum the year data cumulative and then slice to select every fourth row 
# (corresponding to total for the year)

# First gather into long format
depdata <- gather(depdata, state,qtr_departures,2:10) 

# Now do the processing
depdata <- depdata %>%
  group_by(year,state) %>%
  mutate(departures = cumsum(qtr_departures)) %>%
  ungroup() %>%
  slice(seq(4,1188, by =4)) %>%
  select(-qtr_departures)

# Join the arrivals and departures data frames
interstateData <- inner_join(arrdata,depdata,by=c("year","state"))

# Save the final cleaned data set 
write.csv(interstateData, file = file.path(cleanDataFolder, "ABS_interstate_clean.csv"), row.names = FALSE)

```

## Estimated resident population for each age and sex

This R chuck cleans the ERP data by sex ang age provided by the ABS. It is 
currently used in the Chlamydia incidence model. 

```{r grouppop}

### Process ERP for sex and age from second and third sheets
erpdata1 <- read_excel(file.path(absFolder,"3101059-ABS_ERP_Age_Sex.xls"),
                      sheet = 2, skip = 9)

erpdata2 <- read_excel(file.path(absFolder,"3101059-ABS_ERP_Age_Sex.xls"),
                      sheet = 3, skip = 9)

# Combine data frames
erpasdata <- cbind(erpdata1,erpdata2[,2:ncol(erpdata2)])

# Setup column names - 101 ages for males, females and all
ages <- as.character(0:100)
colnames(erpasdata) <- c("year",paste("males-",ages,sep = ""),
                         paste("females-",ages,sep = ""),
                         paste("all-",ages,sep = ""))

# Convert first column to years 
erpasdata[,1] <- format(erpasdata[,1],"%Y")
erpasdata$year <- as.numeric(erpasdata$year)

# Save the final cleaned data set 
write.csv(erpasdata, file = file.path(cleanDataFolder, "ABS_ERP_sex_age_clean.csv"), row.names = FALSE)

```

## Estimated age-specific death rates

This R chuck cleans the ABS age-specific death rate data for males and 
females. This is used for estimating the proportion of PLDHIV in each five
year age bin. 

```{r deathrates}

### Process death rates for sex and age from second and third sheets
deathData <- read_excel(file.path(absFolder,"4125-ABS_Gender_Indicators.xls"),
                      sheet = 14, skip = 14)

# Clean up data
years <- 2000:2011
deathData <- deathData[1:19,]
colnames(deathData) <- c("age", paste(c("males", "females"), 
                                      as.character(rep(years, each = 2)), sep = ""))
deathData[1, 1] <- "0"
deathData[nrow(deathData), 1] <- "85+"

# Reshape data
deathData <- gather(deathData, "pop", "deathrate", 2:ncol(deathData))
deathData$pop <- as.character(deathData$pop)

# Split pop column into gender and year and reorder columns
splitPop <- strsplit(deathData$pop, "s")
deathData$sex <- unlist(lapply(splitPop, function(x) x[1]))
deathData$year <- as.numeric(unlist(lapply(splitPop, function(x) x[2])))
deathData <- deathData %>%
  select(-pop) %>%                 # remove pop
  select(-deathrate, everything()) # put deathrate at end

# Save the final cleaned data set 
write.csv(deathData, file = file.path(cleanDataFolder, "ABS_deathrate_sex_age.csv"), row.names = FALSE)

```

