ABS data cleaning 
=================
  
This script is used to load and clean raw demographic data from the 
Australian Burea of Statistica (ABS). These data are primarily used for 
care cascade calculations but can be useful for Australian evaluation models
in general. The resulting cleaned data sets are saved as dataframes in csv 
files in the ~/Evaluation\_Modelling/data/ folder. 

```{r initialization}
# Clear workspace
rm(list=ls()) 

# Libraries for data manipulation
require(readxl)
require(dplyr)
require(tidyr)
require(utils)

# Key folder paths
basePath <- file.path(path.expand("~"),"Research","!Evaluation_Modelling")
absFolder <- file.path(basePath,"data","ABS") # Folder with the raw ABS data
# dataPath <- file.path(basePath,"data","raw")
cleanDataFolder <- file.path(basePath,"data")

# Set working directory
setwd(file.path(basePath,"code"))

# Order of states used by ABS
states <- c("nsw","vic","qld","sa","wa","tas","nt","act","all")

# Primary script parameters
# dataYear <- 2013

```

## Proportion of people who leave Australia

The final R chuck extracts net overseas migration and estimated resident 
popultion data. It is primarily used to estimate the proportion of the 
which moves overseas each year population 

```{r migration}
# This script cleans the ABS data into the format we want. The raw data is 
# always on the second sheet of the .xls data file.

### Process ERP data

# Read in second sheet skipping most of the bumph
erpdata <- read_excel(file.path(absFolder,"310104-ABS_ERP_State_Overall.xls"),
                      sheet = 2, skip = 9)

# Delete unnecessary columns (keeping the last nine) 
ncolumns <- ncol(erpdata)
erpdata <- erpdata[,c(1,20:28)]

# Convert first column to years 
erpdata[,1] <- format(erpdata[,1],"%Y")

# Rename columns
colnames(erpdata) <- c("year",states)

# Remove unnecessary rows (keeping the final quarter data for each year)
# Starts in second quarter 1981 so selected ever third row
nrows <- nrow(erpdata)
erpdata <- slice(erpdata,seq(3,135,by = 4))

# Reshape into long format
erpdata <- gather(erpdata,state,erp,2:10)
erpdata <- arrange(erpdata,year)

# Convert to correct data types
erpdata$year <- as.numeric(erpdata$year)
erpdata$state <- as.character(erpdata$state)

### Process departure data - Have to read in 9 sheets

# Need to specify years first
years <- 2004:2012

# Initilize results data frame
nomdata <- data.frame(year = as.numeric(),
                      state = as.numeric(),
                      arrivales = as.numeric(),
                      departures = as.numeric(),
                      nom = as.numeric()
)

# Loop through sheets extracting data we want
for (ii in seq(along=years)+1) {
  
  # Read in sheet from file
  tempdata <- read_excel(file.path(absFolder,"34120do007_201314-ABS_Migration_NOM.xls"),
                         sheet = ii, skip = 5)
  
  # Replace annoying column names
  colnames(tempdata) <- c("state","grouping","category","arrivals","departures","nom")
  
  # Select the data we want
  tempdata <- tempdata %>%
    select(-category) %>%
    filter(grouping == "Total")
  
  # General tiding up and add year of data to front
  tempdata$state <- states
  tempdata <- select(tempdata,-grouping) 
  tempdata$year <- rep(years[ii-1],nrow(tempdata)) 
  tempdata <- select(tempdata,year,everything())  # Put year first
  
  # Append to final results
  nomdata <- rbind(nomdata,tempdata)
}

# Order nomdata by state
nomdata <- arrange(nomdata,year)

# Join th nomdata and erpdata sets. 
migrationData <- inner_join(erpdata,nomdata,by=c("year","state"))

# Save the final cleaned data set 
write.csv(migrationData, file = file.path(cleanDataFolder, "ABS_migration_clean.csv"), row.names = FALSE)

```

## Proportion of people who move interstate

```{r interstate}

### Process arrivals read in second sheet
arrdata <- read_excel(file.path(absFolder,"3101016a-ABS_Interstate_Arrivals.xls"),
                      sheet = 2, skip = 9)

# Convert first column to years 
arrdata[,1] <- format(arrdata[,1],"%Y")

# Convert colnames
colnames(arrdata) <- c("year",states)

# Remove first three rows as they form an incomplete year
nrows <- nrow(arrdata)
arrdata <- arrdata[4:nrows,]

# Sum the year data cumulative and then slice to select every fourth row 
# (corresponding to total for the year)

# First gather into long format
arrdata <- gather(arrdata, state,qtr_arrivals,2:10) 

# Now do the processing
arrdata <- arrdata %>%
  group_by(year,state) %>%
  mutate(arrivals = cumsum(qtr_arrivals)) %>%
  ungroup() %>%
  slice(seq(4,1188, by =4)) %>%
  select(-qtr_arrivals)
 

### Process departures
depdata <- read_excel(file.path(absFolder,"3101016b-ABS_Interstate_Departures.xls"),
                      sheet = 2, skip = 9)

# Convert first column to years 
depdata[,1] <- format(depdata[,1],"%Y")

# Convert colnames
colnames(depdata) <- c("year",states)

# Remove first three rows as they form an incomplete year
nrows <- nrow(depdata)
depdata <- depdata[4:nrows,]

# Sum the year data cumulative and then slice to select every fourth row 
# (corresponding to total for the year)

# First gather into long format
depdata <- gather(depdata, state,qtr_departures,2:10) 

# Now do the processing
depdata <- depdata %>%
  group_by(year,state) %>%
  mutate(departures = cumsum(qtr_departures)) %>%
  ungroup() %>%
  slice(seq(4,1188, by =4)) %>%
  select(-qtr_departures)

# Join the arrivals and departures data frames
interstateData <- inner_join(arrdata,depdata,by=c("year","state"))

# Save the final cleaned data set 
write.csv(interstateData, file = file.path(cleanDataFolder, "ABS_interstate_clean.csv"), row.names = FALSE)

```

## Estimated resident population for each age and sex

This R chuck cleans the ERP data by sex ang age provided by the ABS. It is 
currently used in the Chlamydia incidence model. 

```{r grouppop}

### Process ERP for sex and age from second and third sheets
erpdata1 <- read_excel(file.path(absFolder,"3101059-ABS_ERP_Age_Sex.xls"),
                      sheet = 2, skip = 9)

erpdata2 <- read_excel(file.path(absFolder,"3101059-ABS_ERP_Age_Sex.xls"),
                      sheet = 3, skip = 9)

# Combine data frames
erpasdata <- cbind(erpdata1,erpdata2[,2:ncol(erpdata2)])

# Setup column names - 101 ages for males, females and all
ages <- as.character(0:100)
colnames(erpasdata) <- c("year",paste("males-",ages,sep = ""),
                         paste("females-",ages,sep = ""),
                         paste("all-",ages,sep = ""))

# Convert first column to years 
erpasdata[,1] <- format(erpasdata[,1],"%Y")
erpasdata$year <- as.numeric(erpasdata$year)

# Save the final cleaned data set 
write.csv(erpasdata, file = file.path(cleanDataFolder, "ABS_ERP_sex_age_clean.csv"), row.names = FALSE)

```
