---
title: "HIV Cascade Notifications Data"
author: "Richard T. Gray"
date: "Wednesday, July 03, 2015"
output:
  word_document:
    highlight: espresso
---

This R Markdown script is used to process raw HIV notifications files 
produced by the Surveillance, Evaluation Research Program at the Kirby 
Institute, Faculty of Medicine, UNSW. This processing is to produce a 
de-identified data file (in CSV format) with only the necessary categories 
for evaluation purposes.

Raw notifications files MUST remain in the working directory of this 
script. The processed and cleaned files, however, can be copied by Kirby 
staff for further processing.  

This document is written in dynamic format using R markdown v2 within R 
studio 0.98.1091 (using `r substr(R.Version()$version.string,1,15)`). To 
reproduce the cleaned file the user needs to specify the year of the 
notifications data for processing and run this script in Rstudio. As the 
raw notifications data is stored as a Stata .dta data file (currently 
produced by Stata Version 12) users need the 'foreign' R package installed 
to load the file (For Stata 13 the 'haven' R package should be installed 
instead). The final CSV data file is processed using the dplyr library.

To produce a Word document description of this script requires the  the 
installation of the R Markdown v2, knitr, and pandoc packages. <!--- To 
produce the Word Document requires saving in a local directory and setting 
eval = FALSE in the loaddata chunk ---> 


### Script Setup

The following code gets things set up for processing and loads the required
libraries.

```{r setup, message = FALSE, tidy = TRUE}
# Clear workspace
rm(list=ls()) 

# Set working directory
setwd(file.path("/","SVR-NAS","Public","SERP","Cascades","HIV"))

# Load libraries for data manipulation - they are installed if not already
# done so

if (!require(foreign)) {    
  # Try to install package
  install.packages("foreign")
  
  # Try to reload package
  if (!require(foreign)) {
    stop("could not install foreign")
  }
}

if (!require(readstata13)) {    
  # Try to install package
  install.packages("readstata13")
  
  # Try to reload package
  if (!require(foreign)) {
    stop("could not install readstata13")
  }
}

```

### User specified year for processing

The user needs to specify the HIV notifications year. This year is used to 
specify the input and output files. 

```{r User specifications, tidy = TRUE}

# User specified year
yearAnalysis <- 2015

```


```{r files, tidy = TRUE}
# Raw notifications data - MUST REMAIN PRIVATE
rawDataFile <- file.path(getwd(),"data", "National Registry Data",   
                         paste("rawHIVnotifications-",toString(yearAnalysis), 
                               ".dta",sep = ""))

# De-identified output file name
cleanDataFile <- file.path(getwd(),"data",
                           paste("cascadeHIVnotifications-",
                                 toString(yearAnalysis),".csv",
                                                 sep = ""))

cleanDataFileBP <- file.path(getwd(),"data",
                             paste("backprojectionHIVnotifications-",                                                 toString(yearAnalysis),".csv",
                                                 sep = ""))

```

### Load and process the raw HIV notifications file

The raw DTA file is loaded into the work space and the following 
key identifying variables are removed:

* nathivno
* famname
* givename  

The variables required for evaluation modelling are then extracted and 
saved in a de-identified CSV file. <!--- To get the Word Document you need 
to set eval = FALSE in each chunk --->

```{r loaddata, tidy = TRUE} 
#eval = FALSE when running R markdown

# Load the raw data file
if (yearAnalysis < 2015) {
  rawData <- read.dta(rawDataFile) 
} else {
  rawData <- read.dta13(rawDataFile) 
  colnames(rawData) <- tolower(colnames(rawData))
  
  # Sort out indigenous variable
  # rawData$indigenous <- as.numeric(rawData$indigenous)
  # rawData[rawData$indigenous == 1, ]$indigenous <- 0
  # rawData[rawData$indigenous == 2, ]$indigenous <- 1
  # rawData[rawData$indigenous == 3, ]$indigenous <- 9
}

# Select the variables required and create processed data
vars <- c("state", "sex", "dob", "cob", "rob", "datehiv", "age_hiv", 
          "dateaids", "cd4count", "expcode", "partnercob", "dateneg",
          "dateindet", "dateill", "datedeath",
          "causeofdeath", "previ_diag_overseas",
          "country_prev_diag", "cd4_date", "yearhiv")

if (yearAnalysis < 2015) {
  vars <- c(vars,"aborig")
} else {
  vars <- c(vars, "indigenous")
}

processData <- rawData[,vars]

# Save processed data
write.csv(processData, file = cleanDataFile, row.names = FALSE)

### Back projection file 

# Select the variables we need for back projection
varsBP <- c("postcode", "sex", "dob", "cob", "datehiv", "dateaids",
            "cd4count", "expcode", "dateneg", "dateill", 
            "dateindet", "previ_diag_overseas") 
  
processDataBP <- rawData[,varsBP]

# Save back projection data
write.csv(processDataBP, file = cleanDataFileBP, row.names = FALSE)

```

### Clean-up 

Finally clean up work space so nothing is accidentally saved in the wrong place.

```{r cleanup, tidy = TRUE}
# Clear workspace
rm(list=ls())
```
