MBS data cleaning 
=================
  
This script is used to load and clean raw service usage data from the 
Australian Medicare Benefits Scheme (MBS). These data are primarily used for 
care cascade calculations but can be useful for Australian evaluation models
in general. The resulting cleaned data sets are saved as dataframes in csv 
files in the ~/Evaluation\_Modelling/data/ folder. 

```{r initialization}
# Clear workspace
rm(list=ls()) 

# Libraries for data manipulation
require(readxl)
require(dplyr)
require(tidyr)
require(utils)

# Key folder paths
basePath <- file.path(path.expand("~"),"Research","!Evaluation_Modelling")
mbsFolder <- file.path(basePath,"data","MBS") # Folder with the raw MBS data
cleanDataFolder <- file.path(basePath,"data")

# Set working directory
setwd(file.path(basePath,"code"))

# Order of states used by MBS
states <- c("nsw","vic","qld","sa","wa","tas","act","nt","all")

# Primary script parameters
startyear <- 2003
endyear <- 2014

# File tag
mbsFile <- "MBS_Data_"

```

### HIV testing data 

This R chuck extracts the CD4 and viral load testing numbers from the yearly item number data. 

```{r hivtesting}
# The script cleans the MBS HIV testing data into the format we want. 

# Item numbers for HIV testing
items <- c(69378, 69381, 71139)

# Number of rows to extract for each item - manually entered as each item is 
# slightly different
extractRows <- c(30, 30, 36) # 
skipRows <- c(5, 5, 5)# 
numAges <- c(10, 10, 12) #

sexs <- c("female", "male", "all")

# Loop through each item and clean up the data
for (ii in seq(along = items)) {
  # Specify item
  tempItem <- items[ii]
  
  # Specify folder
  tempFolder <- file.path(mbsFolder, toString(tempItem))
  
  # Initialize data frame for storage
  tempDf <- data.frame(year = integer(), sex = character(), age = character(),
                       nsw = integer(), vic = integer(), qld = integer(), 
                       sa = integer(), wa = integer(), tas = integer(),
                       act = integer(), nt = integer(), all = integer())
  
  # Loop through years and extract data
  for (jj in startyear:endyear) {
    # Load file 
    tempData <- read_excel(file.path(tempFolder,
                                     paste(mbsFile, toString(tempItem), "-", 
                                           toString(jj), ".xls", sep = "")),
                           skip = skipRows[ii],
                           col_types = c("text", "text", "numeric", "numeric",
                                        "numeric", "numeric", "numeric", 
                                        "numeric", "numeric", "numeric",
                                        "numeric"))
    
    # Delete unnecesary rows
    tempData <- tempData[1:extractRows[ii],]
    
    # Change column names
    colnames(tempData) <- c("sex","age",states)
    
    # Fill in missing sex values
    for (kk in 1:3) {
       tempData[((kk-1)*numAges[ii]+1):(kk*numAges[ii]),]$sex <- sexs[kk]
    } 
    
    # Clean up age groups
    tempData$age <- tolower(tempData$age)
    tempData[tempData$age == "total",]$age <- "all"
    
    # For item 71139 fix 5-14 age group
    if (tempItem == 71139) {
      tempData[c(2,14,26),]$age <- as.character("5-14")
      # Annoyingly excel may still show this as "May-14" but read.csv 
      # reads in the correct value
    }
    
    # Add year and reorder
    tempData$year <- jj
    tempData <- select(tempData, year, everything())
    
    # Bind each years data 
    tempDf <- rbind(tempDf,tempData)
    
    
  } 
  
  # Write final data frame to file
  write.csv(tempDf, file = file.path(cleanDataFolder, 
                                       paste(mbsFile, toString(tempItem), 
                                             "-clean.csv", sep = "")), row.names = FALSE)
}

```

